---
title: "Geographic Center"
output: html_notebook
---

This notebook shows one method to calculate the center of gravity between a number of points. I have created some dummy data with a separate script.

```{r, echo=TRUE}
# read in the dummy data
library(tidyverse)
library(ggmap)
library(geosphere)
vp <- read.csv("dummy_geo_data.csv")


vp
```


The method to calculate the median distance between center and the points uses the Vincente method assuming the earth is a spheroid:

```{r}
calc_median <- function(center, places){
  distances <- numeric()
  for (i in 1:nrow(places)){
    distances <- c(distances, distVincentySphere(center, c(places$lon[i],places$lat[i])))  
  }
  
  return(round(median(distances)/1000,2))
}
```

Next, I define a function to calculate the center of gravity between various points. The lat / lon coordinates are transposed to carthesian three-dimensional coordinates; then their respective average is calculated; finally, these values are transposed back to lat / lon degrees:

```{r}
calc_center <- function(places){
  # Convert Lat / Lon from degree to radians
  places$latr <- places$lat * pi / 180
  places$lonr <- places$lon * pi / 180
  
  # Convert latr / lonr to Cartesian coordinates for fall locations.
  places$x = cos(places$latr) * cos(places$lonr)
  places$y = cos(places$latr) * sin(places$lonr)
  places$z = sin(places$latr)
  
  # Compute averages for x, y, z coordinates
  x = mean(places$x)
  y = mean(places$y)
  z = mean(places$z)
  
  # Convert average cartesian coordinate back to lat / lon
  Lon = atan2(y, x)
  Hyp = sqrt(x * x + y * y)
  Lat = atan2(z, Hyp)
  
  # Convert Lat / Lon to degrees.
  Lat = Lat * 180 / pi
  Lon = Lon * 180 / pi 
  
  # calculate the median distance from the centroid to the points
  center <- c(Lon, Lat) # SIC! order is reversed!
  median_dist <- calc_median(center, places)
  
  return(c(Lat, Lon, median_dist))
}
```



Then I calculate the center and median distance for each agent in the dummy data set:




```{r}
centers <- data.frame(agent_id=numeric(), lat=numeric(), lon=numeric(), median_dist=numeric())

for (i in unique(vp$agent_id)){
  centers = rbind(centers,c(i,calc_center(filter(vp, agent_id==i))))
}

colnames(centers) <- c("agent_id", "lat", "lon", "median_dist")

centers
```


Finally, these dat can be visualized using ggmap:


```{r}
loc <- "Germany"
map <- qmap(loc, zoom =6)

g <- map + 
  geom_point(aes(x = lon, y = lat, colour = median_dist), size = 2, alpha= 0.5, data = centers)
```




